---
title: "Coding Assignment 2"
author: "Team 9"
date: "Due: 2024-10-13 23:59"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
#Put any packages you need here

library(readxl)

sessionInfo()
knitr::opts_chunk$set(echo = TRUE)


```

A Florida health insurance company wants to predict annual claims for individual clients. The company pulls a random sample of 100 customers. The owner wishes to charge an actuarially fair premium to ensure a normal rate of return. The owner collects all of their current customer’s health care expenses from the last year and compares them with what is known about each customer’s plan. 

The data on the 100 customers in the sample is as follows:

-	Charges: Total medical expenses for a particular insurance plan (in dollars)
-	Age: Age of the primary beneficiary
-	BMI: Primary beneficiary’s body mass index (kg/m2)
-	Female: Primary beneficiary’s birth sex (0 = Male, 1 = Female)
-	Children: Number of children covered by health insurance plan (includes other dependents as well)
-	Smoker: Indicator if primary beneficiary is a smoker (0 = non-smoker, 1 = smoker)
-	Cities: Dummy variables for each city with the default being Sanford

Answer the following questions using complete sentences and attach all output, plots, etc. within this report.


```{r dataset, include=FALSE}

insurance <- read.csv("../CodingAssignment02/Insurance_Data_Group9.csv")
```


## Question 1

Generate this regression equation:

$$Charges = \beta_{0}+\beta_{1}*Age+\beta_{2}*BMI+\beta_{3}*Children + \beta_{4}*Female + \beta_{5}*Smoker + \beta_{6}*WinterSprings + \beta_{7}*WinterPark + \beta_{8}*Oviedo$$

```{r q1}

# Regression equation
model <- lm(Charges ~ Age + BMI + Children + Female + Smoker + WinterSprings + WinterPark + Oviedo, data = insurance)

# Display the summary of the model
regsummary <- summary(model)
regsummary
#summ(model)

# Get info for coefficients and p-values
coefficients <- regsummary$coefficients

# Grab p-values from the 4th column
p_values <- coefficients[, 4]

# Find significant variables at alpha = 0.2
alpha <- 0.2
sigvar <- names(p_values[p_values <= alpha])

#Remove Intercept from significant variables
sigvar <- sigvar[sigvar != "(Intercept)"]

# Print significant variables
print(sigvar)
```

Which independent variables test significant in a two-tailed test? Use $\alpha = 0.2$

Age, BMI, Children, Smoker, and WinterPark all tested significant because p < alpha.


## Question 2

Interpret the coefficients of all the independent variables that tested significant.

For each year of age, there is an estimated additional 251.26 in charges. The margin of error for the age estimation is +/- 59. For each additional point of BMI, there is an estimated additional 394.14 in charges. The margin of error the BMI estimation is +/- 130. For each additional Child covered by the insurance plan there is an estimated additional 1075.41 in charges. The margin of error for the Child estimation is +/- 695. If the primary beneficiary is a smoker there an estimated additional charge of 24707.50. The margin of error for the smoker estimation is +/- 1747. If the city is Winter Park, there is an expected reduction of charges to the tune of -3937.16. The margin of error for the Winter Park estimation is +/- 2052.

## Question 3

The insurance agent returns with another client:

- Age: 35
- BMI: 26
- Children: 1
- Gender: Female
- Smoker: No
- City: Sanford

Using the regression equation above, predict the amount of medical expenses associated with this policy. (Provide a 95% confidence interval as well)

```{r q3}

# Load necessary library
library(stats)

# insurance <- read.csv('insurance.csv')

# Fit the linear regression model
model <- lm(Charges ~ Age + BMI + Children + Female + Smoker + WinterSprings + WinterPark + Oviedo, data = insurance)

# Display the summary of the model to see the coefficients
summary(model)

# New client's information
new_client <- data.frame(
  Age = 35,
  BMI = 26,
  Children = 1,
  Female = 1,   # Female = 1
  Smoker = 0,   # Smoker = No
  WinterSprings = 0,  # Not in WinterSprings
  WinterPark = 0,     # Not in WinterPark
  Oviedo = 0          # Not in Oviedo
)

# Predicting medical expenses with confidence interval
prediction <- predict(model, newdata = new_client, interval = "confidence", level = 0.95)

# Print the prediction and confidence interval
print(prediction)



```
This is a summary of the regression model that includes coefficients showing how much each factor contributes to predicting medical expenses.

The predicted value for this client is $5,575 with a 95% confidence interval of [$2,606.52,$8,544.95]. 

Fit= Charges 
lwr= lower bound
upr= upper bound


## Question 4

Compare the model from question 1, to a model that does not include the group of city dummies. Does including the cities reduce the SSE by a statistically significant amount? (check the partial F-test for significance of the dummies as a group). Use $\alpha = 0.05$

```{r q4}

reduced_insurance <- insurance[,-(7:9)] # drops columns 7 through 9 from original insurance data, eliminating city dummy variables

model_reduced <- lm(Charges ~. , data = reduced_insurance) #creates linear model of data based on reduced insurance model

summary(model_reduced) #creates summary of reduced insurance model

anova(model_reduced, model) #creates comparison of reduced model versus original model including all data

```


